Rust for Scientific Computing

### 2025-09-13 paper/main.tex
continue writing introduction

% keywords can be removed
\keywords{Rust \and HPC \and High-Performance Computing \and Programming}

%###################################################################################################
\section{Introduction}
\label{section:introduction}

General Remarks about languages
\begin{enumerate}
    \item Programming languages allow us to express machine instructions in human-readable code.
    \item Depend on use-case
    \item Translate human-interpretable code into machine instructions
    \item tradeoff between high-level view (abstraction) and performance (low level)
    \item C++ has "zero-cost" abstractions
\end{enumerate}

@@

>>>>

\printinunitsof{in}\prntlen{\columnwidth}

% keywords can be removed
\keywords{Rust \and HPC \and Scientific Computing \and Programming}

%###################################################################################################
\section{Introduction}

% General Remarks about languages
% \begin{enumerate}
%     \item Programming languages allow us to express machine instructions in human-readable code.
%     \item Depend on use-case
%     \item Translate human-interpretable code into machine instructions
%     \item tradeoff between high-level view (abstraction) and performance (low level)
%     \item C++ has "zero-cost" abstractions
% \end{enumerate}

Programming languages sit at the heart of all things computation.
They allow users to express higher-level instructions in a human-readable fashion such that this
code can be translated into machine-level instructions by either compiling or interpreting it.
Various classes of programming languages have emerged that serve distinct purposes.
Fundamentally, they aim to descibe problems at various levels of depth which allows humans to
interpret and reason about the executed instructions and therefore express their desired actions
more clearly.

@@

Many languages have been devised to target specific realms which have emerged over the past decades.
They span across a range of abstraction levels, ranging from low-level languages to high-level ones.
The low-level operations level mostly deals with instructions on a level close to the actual
hardware such as memory allocation, pointer arithmetic or by using assembly instructions directly.
Meanwhile, high-level operations usually combine many lower-level ones to perform a single task.
An example for a high-level operation would be the addition of two matrices which might allocate the
required memory, fill it according to the given values and then call the appropriate math function
to pairwise add their entries.
Languages which can operate effectively on a lower level often bear the most promise for maximum
performance which is crucial in any scientific application.
These languages are mostly compiled using an optimizing compiler which enables many of their
performance achievements.

This mini-review concerns itself with the Rust programming language and aims to provide an overview
of the field of scientific computing with the Rust ecosystem.
Other systems programming languages  such as C, C++ or Fortran have dominated the field over the
past decades.
Most notably, C++ has been influential in establishing a model of "zero cost abstractions" where a
phrasing of the problem from a higher point of view does not involve any additional execution cost
but is handled by the compiler.
These and many other principles have allowed for a plethora of
